Predicting instabilities of a tuneable ring laser with an iterative map model

### 2021-06-14 Papers/Journal/Tuneable_Lasers.tex
Made final grammar changes

Several effects arise within the cavity due to the interplay of dispersion, modulation, and nonlinearity~\cite{bohun2015, coen1997, lapre2019, meng2020, oktem2010, shao2019, woodward2018}. One effect, wave breaking, causes the leading edge of a pulse to redshift, and the trailing edge to blueshift (the effect reverses in the anomalous dispersion regime) to the point a shock develops~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The frequency shift causes the pulse to become more rectangular in the frequency domain, with a linear chirp over most of the pulse, saturating at the leading and trailing edge. In optics, wave breaking arises from self-phase modulation (SPM), which is a direct effect of the Kerr nonlinearity, causing the pulse to interfere with itself. New frequencies are then generated such that a rectangular profile is induced~\cite{agrawal2013, woodward2018}. Outside of optics, wave breaking can also occur in other areas with nonlinear waves and dispersive media such as plasma physics, heat propagation, transmission lines, and fluid dynamics~\cite{coen1997, rothenberg1989b}. Within a ring laser cavity, the combination of dispersion, modulation, and nonlinearity can also lead to instabilities that degrade the pulse. Wave breaking and instabilities can become parasitic, leading to an unstable and unsustainable pulse. Understanding the rich landscape of the parameter space is important for determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

@@

The GNLSE, \eqref{eq:nlse}, has been applied in nonlinear optics~\cite{agrawal2013} and fibre optic communications~\cite{agrawal2002}. However, \eqref{eq:nlse} is regularly generalized even further to include a modulation term. The resulting equation is referred to as the master equation of mode-locking (see \cite{haus1975, haus1984, haus2000, tamura1996, usechak2005}), or sometimes as the Ginzburg--Landau equation. Owing to the nonlinear term in \eqref{eq:nlse} and the often non-trivial form of $g(A)$, as well as the possible addition of a modulation term, no analytic solution is known to the master equation or to \eqref{eq:nlse}. A comprehensive description of the theory and history of mode-locking lasers is presented in Haus~\cite{haus2000}. While the GNLSE well represents the dynamics in a single optical fibre, \eqref{eq:nlse} is not entirely representative of the underlying physics within a laser cavity such as Fig. \ref{fig:cavity}. The derivation of \eqref{eq:nlse} assumes that each process affects the pulse continuously within the cavity. The assumption that a single GNLSE applies to the whole laser cavity breaks down if the pulse undergoes large changes during any one round trip. Often, this assumption does not hold because of the specialized, discrete components.

@@

\section{Iterative Map Model}
\label{sec:model}
We derive our own iterative map model for the evolution of a pulse through a laser cavity by solving a reduced GNLSE for each component in the cavity---except the modulator, which we consider to be applied externally by an electro-optic modulator. Starting with the gain component, we reduce \eqref{eq:nlse} to 
\begin{align}
	\pdiff{A}{z} &= \frac{1}{2}g(A)A,
	\label{eq:gainexample}
\end{align}
by assuming all other effects are negligible. Once again, our assumption is justified by the localization of each effect to its corresponding component, allowing us to simplify \eqref{eq:nlse} to \eqref{eq:gainexample} in the gain fibre. To proceed, we assume the gain has the form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},
	\label{eq:gainform}
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, defined as
\begin{align}
	E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
and $\Es$ is the energy that the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}, which depends on the power of the pump laser. With the choice of $g(A)$ given in \eqref{eq:gainform}, \eqref{eq:gainexample} becomes
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
	\label{eq:pdiffgain}
\end{align}
Multiplying \eqref{eq:pdiffgain} by the complex conjugate of $A$, we have
\begin{align}
	\overline{A} \pdiff{A}{z} &= \frac{g_0}{2 \left( 1 + E / \Es \right)} |A|^2.
	\label{eq:pdiffgain2}
\end{align}
By adding \eqref{eq:pdiffgain2} and the complex conjugate of \eqref{eq:pdiffgain2}, and then integrating over $T$, we have
\begin{align}
	\diff{E}{z} &= \frac{g_0 E}{1 + E / \Es}.
	\label{eq:energyde}
\end{align}
By assuming the energy entering the gain fibre is $E_g$, and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we integrate \eqref{eq:energyde} to find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E_g}{\Es} \exp \left( \frac{E_g}{\Es} \right) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Moreover, as the gain will amplify the pulse without changing the shape of the envelope, the pulse after travelling through the gain fibre can be expressed as
\begin{align}
	G(A) &= \mathcal{G} A.
	\label{eq:ampfact}
\end{align}
By \eqref{eq:energy}, \eqref{eq:engain} and \eqref{eq:ampfact},
\begin{align}
	\mathcal{G}^2 E_g &= E_\text{out},
\end{align}
therefore,
\begin{align}
	G(A) &= \left( \frac{E_\text{out}}{E_g} \right)^{1/2} A \\
	&=	\left( \frac{\Es}{E_g} W \left( \frac{E_g}{\Es} \exp \left( \frac{E_g}{\Es} \right) \exp( g_0 L_g ) \right) \right)^{1/2} A \label{eq:dimgain}
\end{align}
for the amplification of the pulse after the Er-doped gain fibre.

Repeating the same procedure for the nonlinearity, loss, and dispersion components yields
\begin{align}
	\pdiff{A}{z} &= i \gamma |A|^2 A, \label{eq:nlde} \\
	\pdiff{A}{z} &= - \alpha A,  \label{eq:lossde} \\
	\pdiff{A}{z} &= - i \frac{\beta_2}{2} \pdiff[2]{A}{T} \label{eq:dispde}
\end{align}
as the reduced versions of \eqref{eq:nlse}, respectively. Solving \eqref{eq:nlde} over a length of fibre $L_f$, we find the effect of the fibre nonlinearity is
\begin{align}
	F(A) &= A \exp \left( i \gamma |A|^2 L_f \right). \label{eq:dimnl}
\end{align}
Similarly, the effect of the loss is
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \label{eq:dimloss}
\end{align}
where $L_T$ is the total length of fibre in the cavity, $R$ is the reflectivity of the optical coupler, and the $(1 - R)$ term  is the loss to the output of the laser. The effect of dispersion is obtained by solving \eqref{eq:dispde} over the `length' of the dispersive medium, $L_D$,
\begin{align}
	D(A) &= \FTi{\exp( i \omega^2 \beta_2 L_D / 2 ) \FT{A}}, \label{eq:dimdisp}
\end{align}
where, $\mathcal{F}$ and $\mathcal{F}^{-1}$ denote the Fourier transform and its inverse.

Finally, we consider the modulation. We assume the pulse is modulated by an electro-optic modulator. For simplicity, the form is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of modulation. This assumption is not restrictive. The results apply to any square integrable modulation function through a result of Calcaterra and Boldt~\cite{calcaterra2008a} who prove that linear combinations of translations of a single Gaussian, $\textrm{e}^{-x^2}$, are dense in $L^2(\mathbb{R})$. Any continuous modulation function with compact support would fall into this class. Additionally, using a saturable absorber in place of an electro-optic modulator is common in practice. In this case, we can exchange \eqref{eq:modform} for~\cite{lapre2019, meng2020, oktem2010, woodward2018}
\begin{align}
	S(A) &= \left( 1 - \frac{q_0}{1 + |A|^2 / P_\text{sat}} \right) A,
\end{align}
where $q_0$ is the modulation depth, and $P_\text{sat}$ is the saturation power.

@@

To further justify the decoupling of the GNSLE into an iterative map, \eqref{eq:nlse} can be written in a dimensionless form where the characteristic time, length, and amplitude are taken to be $T_M$, $L$, and $P_0^{1/2}$, respectively. Letting
\begin{align}
	T &= T_M \widetilde{T}, & z &= L \widetilde{z}, & A &= P_0^{1/2} \widetilde{A},
\end{align}
\eqref{eq:nlse} becomes
\begin{align}
	\pdiff{\widetilde{A}}{\widetilde{z}} = \frac{-i \, \textrm{sgn}(\beta_2) L}{L_\textrm{disp}} \pdiff[2]{\widetilde{A}}{\widetilde{T}} + i \frac{L}{L_\textrm{NL}} |\widetilde{A}|^2 \widetilde{A} + \frac{1}{2}g_0 L \, \widetilde{g}(\widetilde{A}) \widetilde{A} - \alpha L \widetilde{A},
	\label{eq:ndgnlse}
\end{align}
where
\begin{equation}
	\begin{gathered}
		\widetilde{g}(\widetilde{A}) = \left( 1 + \frac{P_0 T_M}{\Es} \int_{-\infty}^\infty |\widetilde{A}|^2 \, \df \widetilde{T} \right)^{-1}, \\
		L_\textrm{disp} = \frac{T_M^2}{|\beta_2|}, \hspace{2cm} L_\textrm{NL} = \frac{1}{\gamma P_0}.
	\end{gathered}
\end{equation}
The dispersion length $L_\textrm{disp}$, and the nonlinear length $L_\textrm{NL}$, are identified in~\cite{agrawal2013} as length scales over which either the dispersive, or nonlinear, effects dominate the evolution of a pulse.

@@

We consider the pulse, shown in Fig. \ref{fig:samplegauss}, given by
\begin{align}
	A = \sqrt{P} \exp \left( -(1 + iC) \frac{T^2}{2 \sigma^2} \right) \exp(i \phi_0),
	\label{eq:A0}
\end{align}
where $P$ is the peak power, $C$ is the chirp, $\sigma$ is the half-width of the pulse (at $1 / \textrm{e}$ intensity), and $\phi_0$ is the phase. By computing \eqref{eq:A0} after one round trip, and solving $\mathcal{L}(A) = A \textrm{e}^{i \phi}$, tedious manipulations yield the conditions
\begin{align}
	\frac{\sigma^4}{1 - \sigma^2} &= \left( \sigma^2 + 2 C s^2 \right)^2 + 4 s^4, \label{eq:sigrelation} \\
	\frac{C}{1 - \sigma^2} &= C + 2 \frac{s^2}{\sigma^2} \left( 1 + C^2 \right), \label{eq:chirprelation} \\
	1 &= \frac{W(a E_g \textrm{e}^{E_g})}{E_g} h^2 \left( 1 - \sigma^2 \right)^{1/2} \label{eq:energyrelation}
\end{align}
for \eqref{eq:A0} to be in equilibrium. Using \eqref{eq:sigrelation} and \eqref{eq:chirprelation} we eliminate the chirp, $C$, and find an expression relating $\sigma^2$ and $s$,~namely
\begin{align}
	\sigma^8 + 4 s^4 \sigma^6 - 20 s^4 \sigma^4 + 32 s^4 \sigma^2 - 16 s^4 = 0.
	\label{eq:var}
\end{align}
\eqref{eq:var} has the positive, real, analytic solution
\begin{equation}
	\begin{split}
		\sigma^2 = \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} & \\
		- s^4 - s^2 & \sqrt{4 + s^4}.
		\label{eq:equilvar}
	\end{split}
\end{equation}
We find the chirp
\begin{align}
	C = \frac{\sigma^4}{2 s^2 (2 - \sigma^2)}
	\label{eq:chirp}
\end{align}
by combining \eqref{eq:sigrelation} and \eqref{eq:chirprelation}. We obtain the energy of the pulse entering the gain fibre, $E_g$, from \eqref{eq:energyrelation}:
\begin{align}
	E_g &= \frac{h^2 \left( 1 - \sigma^2 \right)^{1/2} \log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{h^2 \left( 1 - \sigma^2 \right)^{1/2} - 1},
\end{align}
and thus, the equilibrium energy at the optical coupler, is
\begin{align}
	E &= W \left( a E_g \textrm{e}^{E_g} \right) \\
	&= \frac{\log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{1 - h^2 \left( 1 - \sigma^2 \right)^{1/2}}.
	\label{eq:analenergy}
\end{align}
The argument of the logarithm in \eqref{eq:analenergy} gives us the condition
\begin{align}
	a h^2 (1 - \sigma^2)^{1/2} > 1
	\label{eq:energycond}
\end{align}
for the pulse to be sustainable.

@@

The forms of \eqref{eq:equilvar}--\eqref{eq:analpower} make the expressions difficult to work with. However, recall that $s$ is small, thus, we asymptotically expand \eqref{eq:equilvar}--\eqref{eq:analpower} assuming $s \ll 1$, and
\begin{align}
	\sigma^2 &\sim 2s(1 - s) + \bigO{s^3}, \\
	C &\sim 1 - s + \frac{1}{2}s^2 + \bigO{s^3}, \\
	E &\sim \frac{\log (a h^2)}{1 - h^2} - \frac{1}{1 - h^2} \left( 1 + \frac{h^2 \log(a h^2)}{1 - h^2}  \right) s + \bigO{s^2}, \label{eq:maxenergy} \\
	P &\sim \frac{\log(ah^2)}{\sqrt{2 \pi}(1 - h^2)} s^{-1/2} + \bigO{s^{1/2}},
\end{align}
in the limit $s \rightarrow 0$. The expansions in the limit $s \rightarrow \infty$ are given in Appendix \ref{sec:largedisp}.

@@

Also of interest is the energy of the pulse at equilibrium, which we compute numerically by \eqref{eq:energy}. Given our ordering of the components, we compute the energy of the pulse immediately before passing through the optical coupler. Thus, the energy of each pulse emitted by the laser is $(1 - h^2) E$, and $h^2 E$ remains in the cavity. We plot the energy in the $s$--$b$ plane in Fig.~\ref{fig:energy}. Unsurprisingly, we find the same sharp boundary as we saw in Fig.~\ref{fig:error}. In the unstable region (upper-left) the energy is small. The pulse is not able to foster prominent modes in this region, and is unsustainable, thus, the intensity is low across the entire pulse. In the stable region (lower-right) we find the energy decays smoothly as both $b$ and $s$ increase, with the contours being hyperbolic-esque. As $s$ increases, the tails of the pulse increase in length due to the dispersion, and more energy is lost due to modulation. As $b$ increases, our choice of modulation function helps suppress the additional frequencies generated by SPM, ensuring the pulse remains band-limited. We find a more energetic laser requires weaker dispersion, however, the nonlinearity must be correspondingly low, otherwise the pulse will succumb to the instability. Moreover, the maximum energy, which occurs in the limit where both $s$ and $b$ go to zero, is 2.5487. With our choice of parameters, \eqref{eq:maxenergy} predicts an equilibrium energy of 2.5535 at the origin, confirming the usefulness of the approximation. Moreover, we show in Fig. \ref{fig:energytime} the energy variation of four pulses during the 500 round trips. In particular, we plot two pulses in the stable region (purple and blue) and two pulses in the unstable region (green and red). Each of the pulses, again, has an initial energy of $E_0 = 0.1$, and after about a dozen round trips the energy is approximately 1.25 in the case of the $s = 0.35$, $b = 6.0$, and approximately 2 in the other three cases. The two examples in the stable region quickly reach equilibrium, and their energy remains constant for the remaining round trips. However, for the examples in the unstable region the pulses lose a considerable amount of their energy and the pulse begins to degrade. The energy of the pulses varies by up to 25\% between consecutive round trips. A pulse in the unstable region undergoes large transformations with each round trip leading to the error we saw in Fig. \ref{fig:error}, as well as the energy variation in Fig. \ref{fig:energytime}.

>>>>

Several effects arise within the cavity due to the interplay of dispersion, modulation, and nonlinearity~\cite{bohun2015, coen1997, lapre2019, meng2020, oktem2010, shao2019, woodward2018}. One effect, wave breaking, causes the leading edge of a pulse to redshift, and the trailing edge to blueshift (the effect reverses in the anomalous dispersion regime) to the point a shock develops~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The frequency shift causes the pulse to become more rectangular in the frequency domain, with a linear chirp over most of the pulse, saturating at the leading and trailing edges. In optics, wave breaking arises from self-phase modulation (SPM), which is a direct effect of the Kerr nonlinearity, causing the pulse to interfere with itself. New frequencies are then generated such that a rectangular profile is induced~\cite{agrawal2013, woodward2018}. Outside of optics, wave breaking can also occur in other areas with nonlinear waves and dispersive media such as plasma physics, heat propagation, transmission lines, and fluid dynamics~\cite{coen1997, rothenberg1989b}. Within a ring laser cavity, the combination of dispersion, modulation, and nonlinearity can also lead to instabilities that degrade the pulse. Wave breaking and instabilities can become parasitic, leading to an unstable and unsustainable pulse. Understanding the rich landscape of the parameter space is important for determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

@@

The GNLSE, \eqref{eq:nlse}, has been applied in nonlinear optics~\cite{agrawal2013} and fibre optic communications~\cite{agrawal2002}. However, \eqref{eq:nlse} is regularly generalized even further to include a modulation term. The resulting equation is referred to as the master equation of mode-locking (see \cite{haus1975, haus1984, haus2000, tamura1996, usechak2005}), or sometimes as the Ginzburg--Landau equation. Owing to the nonlinear term in \eqref{eq:nlse} and the often non-trivial form of $g(A)$, as well as the possible addition of a modulation term, no analytic solution is known to the master equation or to \eqref{eq:nlse}. A comprehensive description of the theory and history of mode-locking lasers is presented in Haus~\cite{haus2000}. While the GNLSE well represents the dynamics in a single optical fibre, \eqref{eq:nlse} is not entirely representative of the underlying physics within a laser cavity such as Fig.~\ref{fig:cavity}. The derivation of \eqref{eq:nlse} assumes that each process affects the pulse continuously within the cavity. The assumption that a single GNLSE applies to the whole laser cavity breaks down if the pulse undergoes large changes during any one round trip. Often, this assumption does not hold because of the specialized, discrete components.

@@

\section{Iterative Map Model}
\label{sec:model}
We derive our own iterative map model for the evolution of a pulse through a laser cavity by solving a reduced GNLSE for each component in the cavity---except the modulator, which we consider to be applied externally by an electro-optic modulator. Starting with the gain component, we reduce \eqref{eq:nlse} to 
\begin{align}
	\pdiff{A}{z} &= \frac{1}{2}g(A)A,
	\label{eq:gainexample}
\end{align}
by assuming all other effects are negligible. Once again, our assumption is justified by the localization of each effect to its corresponding component, allowing us to simplify \eqref{eq:nlse} to \eqref{eq:gainexample} in the gain fibre. To proceed, we assume the gain has the~form
\begin{align}
	g(A) &= \frac{g_0}{1 + E / \Es},
	\label{eq:gainform}
\end{align}
where $g_0$ is a small signal gain, $E$ is the energy of the pulse, defined as
\begin{align}
	E &= \int_{-\infty}^\infty |A|^2 \, \df T,
	\label{eq:energy}
\end{align}
and $\Es$ is the energy that the gain begins to saturate~\cite{haus1984, shtyrina2017, silfvast2004}, which depends on the power of the pump laser. With the choice of $g(A)$ given in \eqref{eq:gainform}, \eqref{eq:gainexample} becomes
\begin{align}
	\pdiff{A}{z} &= \frac{g_0 A}{2 \left( 1 + E / \Es \right)}.
	\label{eq:pdiffgain}
\end{align}
Multiplying \eqref{eq:pdiffgain} by the complex conjugate of $A$, we have
\begin{align}
	\overline{A} \pdiff{A}{z} &= \frac{g_0}{2 \left( 1 + E / \Es \right)} |A|^2.
	\label{eq:pdiffgain2}
\end{align}
By adding \eqref{eq:pdiffgain2} and the complex conjugate of \eqref{eq:pdiffgain2}, and then integrating over $T$, we have
\begin{align}
	\diff{E}{z} &= \frac{g_0 E}{1 + E / \Es}.
	\label{eq:energyde}
\end{align}
By assuming the energy entering the gain fibre is $E_g$, and the energy after travelling through the gain fibre of length $L_g$ is $E_{\text{out}}$, we integrate \eqref{eq:energyde} to find the energy after amplification is
\begin{align}
	\label{eq:engain}
	E_{\text{out}} &= \Es W \left( \frac{E_g}{\Es} \exp \left( \frac{E_g}{\Es} \right) \exp( g_0 L_g ) \right),
\end{align}
where $W$ is the Lambert $W$ function. Moreover, as the gain will amplify the pulse without changing the shape of the envelope, the pulse after travelling through the gain fibre can be expressed~as
\begin{align}
	G(A) &= \mathcal{G} A.
	\label{eq:ampfact}
\end{align}
By \eqref{eq:energy}, \eqref{eq:engain} and \eqref{eq:ampfact},
\begin{align}
	\mathcal{G}^2 E_g &= E_\text{out},
\end{align}
therefore,
\begin{align}
	G(A) &= \left( \frac{E_\text{out}}{E_g} \right)^{1/2} A \\
	&=	\left( \frac{\Es}{E_g} W \left( \frac{E_g}{\Es} \exp \left( \frac{E_g}{\Es} \right) \exp( g_0 L_g ) \right) \right)^{1/2} A \label{eq:dimgain}
\end{align}
for the amplification of the pulse after the Er-doped gain fibre.

Repeating the same procedure for the nonlinearity, loss, and dispersion components yields
\begin{align}
	\pdiff{A}{z} &= i \gamma |A|^2 A, \label{eq:nlde} \\
	\pdiff{A}{z} &= - \alpha A,  \label{eq:lossde} \\
	\pdiff{A}{z} &= - i \frac{\beta_2}{2} \pdiff[2]{A}{T} \label{eq:dispde}
\end{align}
as the reduced versions of \eqref{eq:nlse}, respectively. Solving \eqref{eq:nlde} over a length of fibre $L_f$, we find the effect of the fibre nonlinearity~is
\begin{align}
	F(A) &= A \exp \left( i \gamma |A|^2 L_f \right). \label{eq:dimnl}
\end{align}
Similarly, the effect of the loss is
\begin{align}
	L(A) &= (1 - R) \exp( - \alpha L_T )A, \label{eq:dimloss}
\end{align}
where $L_T$ is the total length of fibre in the cavity, $R$ is the reflectivity of the optical coupler, and the $(1 - R)$ term is the percent not lost to the output of the laser. The effect of dispersion is obtained by solving \eqref{eq:dispde} over the `length' of the dispersive medium, $L_D$,
\begin{align}
	D(A) &= \FTi{\exp( i \omega^2 \beta_2 L_D / 2 ) \FT{A}}, \label{eq:dimdisp}
\end{align}
where, $\mathcal{F}$ and $\mathcal{F}^{-1}$ denote the Fourier transform and its inverse.

Finally, we consider the modulation. We assume the pulse is modulated by an electro-optic modulator. For simplicity, the form is taken as the Gaussian
\begin{align}
	M(A) &= \exp( -T^2 / 2 T_M^2 ) A,
	\label{eq:modform}
\end{align}
where $T_M$ is the characteristic width of modulation. This assumption is not restrictive. The results apply to any square integrable modulation function through a result of Calcaterra and Boldt~\cite{calcaterra2008a} who prove that linear combinations of translations of a single Gaussian, $\exp \left( -x^2 \right)$, are dense in $L^2(\mathbb{R})$. Any continuous modulation function with compact support would fall into this class. Additionally, using a saturable absorber in place of an electro-optic modulator is common in practice. In this case, we can exchange \eqref{eq:modform} for~\cite{lapre2019, meng2020, oktem2010, woodward2018}
\begin{align}
	S(A) &= \left( 1 - \frac{q_0}{1 + |A|^2 / P_\text{sat}} \right) A,
\end{align}
where $q_0$ is the modulation depth, and $P_\text{sat}$ is the saturation power.

@@

To further justify the decoupling of the GNLSE into an iterative map, \eqref{eq:nlse} can be written in a dimensionless form where the characteristic time, length, and amplitude are taken to be $T_M$, $L$, and $P_0^{1/2}$, respectively. Letting
\begin{align}
	T &= T_M \widetilde{T}, & z &= L \widetilde{z}, & A &= P_0^{1/2} \widetilde{A},
\end{align}
\eqref{eq:nlse} becomes
\begin{align}
	\pdiff{\widetilde{A}}{\widetilde{z}} = \frac{-i \, \textrm{sgn}(\beta_2) L}{L_\textrm{disp}} \pdiff[2]{\widetilde{A}}{\widetilde{T}} + i \frac{L}{L_\textrm{NL}} |\widetilde{A}|^2 \widetilde{A} + \frac{1}{2}g_0 L \, \widetilde{g}(\widetilde{A}) \widetilde{A} - \alpha L \widetilde{A},
	\label{eq:ndgnlse}
\end{align}
where
\begin{equation}
	\begin{gathered}
		\widetilde{g}(\widetilde{A}) = \left( 1 + \frac{P_0 T_M}{\Es} \int_{-\infty}^\infty |\widetilde{A}|^2 \, \df \widetilde{T} \right)^{-1}, \\
		L_\textrm{disp} = \frac{T_M^2}{|\beta_2|}, \hspace{2cm} L_\textrm{NL} = \frac{1}{\gamma P_0}.
	\end{gathered}
\end{equation}
The dispersion length $L_\textrm{disp}$, and the nonlinear length $L_\textrm{NL}$, are identified in~\cite{agrawal2013} as length scales over which either the dispersive, or nonlinear, effects dominate the evolution of a pulse.

@@

We consider the pulse, shown in Fig.~\ref{fig:samplegauss}, given by
\begin{align}
	A = \sqrt{P} \exp \left( -(1 + iC) \frac{T^2}{2 \sigma^2} \right) \exp(i \phi_0),
	\label{eq:A0}
\end{align}
where $P$ is the peak power, $C$ is the chirp, $\sigma$ is the half-width of the pulse (at $1 / \textrm{e}$ intensity), and $\phi_0$ is the phase. By computing \eqref{eq:A0} after one round trip, and solving $\mathcal{L}(A) = A \textrm{e}^{i \phi}$, tedious manipulations yield the conditions
\begin{align}
	\frac{\sigma^4}{1 - \sigma^2} &= \left( \sigma^2 + 2 C s^2 \right)^2 + 4 s^4, \label{eq:sigrelation} \\
	\frac{C}{1 - \sigma^2} &= C + 2 \frac{s^2}{\sigma^2} \left( 1 + C^2 \right), \label{eq:chirprelation} \\
	1 &= \frac{W(a E_g \textrm{e}^{E_g})}{E_g} h^2 \left( 1 - \sigma^2 \right)^{1/2} \label{eq:energyrelation}
\end{align}
for \eqref{eq:A0} to be in equilibrium. Using \eqref{eq:sigrelation} and \eqref{eq:chirprelation} we eliminate the chirp, $C$, and find an expression relating $\sigma^2$ and $s$,~namely
\begin{align}
	\sigma^8 + 4 s^4 \sigma^6 - 20 s^4 \sigma^4 + 32 s^4 \sigma^2 - 16 s^4 = 0.
	\label{eq:var}
\end{align}
\eqref{eq:var} has the positive, real, analytic solution
\begin{equation}
	\begin{split}
		\sigma^2 = \sqrt{2} s \left( s^6 + 3s^2 + \sqrt{4 + s^4} \left( 1 + s^4 \right) \right)^{1/2} & \\
		- s^4 - s^2 & \sqrt{4 + s^4}.
		\label{eq:equilvar}
	\end{split}
\end{equation}
We find the chirp
\begin{align}
	C = \frac{\sigma^4}{2 s^2 (2 - \sigma^2)}
	\label{eq:chirp}
\end{align}
by combining \eqref{eq:sigrelation} and \eqref{eq:chirprelation}. We obtain the energy of the pulse entering the gain fibre, $E_g$, from \eqref{eq:energyrelation}:
\begin{align}
	E_g &= \frac{h^2 \left( 1 - \sigma^2 \right)^{1/2} \log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{h^2 \left( 1 - \sigma^2 \right)^{1/2} - 1},
\end{align}
and thus, the equilibrium energy at the optical coupler, is
\begin{align}
	E &= W \left( a E_g \textrm{e}^{E_g} \right) \\
	&= \frac{\log \left( a h^2 \left( 1 - \sigma^2 \right)^{1/2} \right)}{1 - h^2 \left( 1 - \sigma^2 \right)^{1/2}}.
	\label{eq:analenergy}
\end{align}
The argument of the logarithm in \eqref{eq:analenergy} gives us the condition
\begin{align}
	a h^2 (1 - \sigma^2)^{1/2} > 1
	\label{eq:energycond}
\end{align}
for the pulse to be sustainable.

@@

The forms of \eqref{eq:equilvar}--\eqref{eq:analpower} make the expressions difficult to work with. However, recall that $s$ is small, thus, we asymptotically expand \eqref{eq:equilvar}--\eqref{eq:analpower} assuming $s \ll 1$, and find
\begin{align}
	\sigma^2 &\sim 2s(1 - s) + \bigO{s^3}, \\
	C &\sim 1 - s + \frac{1}{2}s^2 + \bigO{s^3}, \\
	E &\sim \frac{\log (a h^2)}{1 - h^2} - \frac{1}{1 - h^2} \left( 1 + \frac{h^2 \log(a h^2)}{1 - h^2}  \right) s + \bigO{s^2}, \label{eq:maxenergy} \\
	P &\sim \frac{\log(ah^2)}{\sqrt{2 \pi}(1 - h^2)} s^{-1/2} + \bigO{s^{1/2}},
\end{align}
in the limit $s \rightarrow 0$. The expansions in the limit $s \rightarrow \infty$ are given in Appendix \ref{sec:largedisp}.

@@

Also of interest is the energy of the pulse at equilibrium, which we compute numerically by \eqref{eq:energy}. Given our ordering of the components, we compute the energy of the pulse immediately before passing through the optical coupler. Thus, the energy of each pulse emitted by the laser is $(1 - h^2) E$, and $h^2 E$ remains in the cavity. We plot the energy in the $s$--$b$ plane in Fig.~\ref{fig:energy}. Unsurprisingly, we find the same sharp boundary as we saw in Fig.~\ref{fig:error}. In the unstable region (upper-left) the energy is small. The pulse is not able to foster prominent modes in this region, and is unsustainable, thus, the intensity is low across the entire pulse. In the stable region (lower-right) we find the energy decays smoothly as both $b$ and $s$ increase, with the contours being hyperbolic-esque. As $s$ increases, the tails of the pulse increase in length due to the dispersion, and more energy is lost due to modulation. As $b$ increases, our choice of modulation function helps suppress the additional frequencies generated by SPM, ensuring the pulse remains band-limited. We find a more energetic laser requires weaker dispersion, however, the nonlinearity must be correspondingly low, otherwise the pulse will succumb to the instability. Moreover, the maximum energy, which occurs in the limit where both $s$ and $b$ go to zero, is 2.5487. With our choice of parameters, \eqref{eq:maxenergy} predicts an equilibrium energy of 2.5535 at the origin, confirming the usefulness of the approximation. Moreover, we show in Fig.~\ref{fig:energytime} the energy variation of four pulses during the 500 round trips. In particular, we plot two pulses in the stable region (purple and blue) and two pulses in the unstable region (green and red). Each of the pulses, again, has an initial energy of $E_0 = 0.1$, and after about a dozen round trips the energy is approximately 1.25 in the case of the $s = 0.35$, $b = 6.0$, and approximately 2 in the other three cases. The two examples in the stable region quickly reach equilibrium, and their energy remains constant for the remaining round trips. However, for the examples in the unstable region the pulses lose a considerable amount of their energy and the pulse begins to degrade. The energy of the pulses varies by up to 25\% between consecutive round trips. A pulse in the unstable region undergoes large transformations with each round trip leading to the error we saw in Fig.~\ref{fig:error}, as well as the energy variation in Fig.~\ref{fig:energytime}.
